Arrows for Parallel Computation

### 2018-01-06 paper/content/utilityFunctions.tex
sorting out references
shorter captions for figs

The |mapArr| combinator (Fig.~\ref{fig:mapArr}) lifts any Arrow |arr a
b| to an Arrow |arr [a] [b]|. The original inspiration was from \citet{programming_with_arrows},
but the definition as then unified with |evalN|.

@@

instance Future BasicFuture a () where
    put _ = put'
    get _ = get'
\end{code}
\caption{The |BasicFuture| type and its |Future| instance for the |Par| Monad and GpH Haskell backends.}
\label{fig:BasicFuture}
\end{figure}

@@

\begin{figure}[thb]
%parMap
\includegraphics[scale=0.7]{images/parMap}
\caption{Schematic depiction of |parMap|.}
\label{fig:parMapImg}

@@

%parMapStream
\includegraphics[scale=0.7]{images/parMapStream}
\caption{Schematic depiction of |parMapStream|.}
\label{fig:parMapStreamImg}

\begin{code}
parMapStream :: (ArrowParallel arr a b conf, ArrowChoice arr, ArrowApply arr) =>
	conf -> ChunkSize -> arr a b -> arr [a] [b]
parMapStream conf chunkSize f = parEvalNLazy conf chunkSize (repeat f)
\end{code}
\caption{Definition of |parMapStream|.}
\label{fig:parMapStream}
\end{figure}

@@

takeEach :: Int -> [a] -> [a]
takeEach n [] = []
takeEach n (x:xs) = x : takeEach n (drop (n-1) xs)
\end{code}
\caption{Definitions of |shuffle|, |unshuffle|, |takeEach|.}
\label{fig:edenshuffleetc}
\end{figure}

@@

rightRotate :: (Arrow arr) => arr [a] [a]
rightRotate = arr $ \list -> case list of
  [] -> []
  xs -> last xs : init xs
\end{code}
\caption{Definitions of |lazy| and |rightRotate|.}
\label{fig:edenlazyrightrotate}
\end{figure}

@@

threetotwo :: (Arrow arr) => arr (a, b, c) (a, (b, c))
threetotwo = arr $ \ ~(a, b, c) -> (a, (b, c))
\end{code}
\caption{Definitions of |lazyzip3|, |uncurry3| and |threetotwo|.}
\label{fig:lazyzip3etc}
\end{figure}

\begin{figure}[h]
\begin{code}
farmChunk :: (ArrowParallel arr a b conf, ArrowParallel arr [a] [b] conf, 
             ArrowChoice arr, ArrowApply arr) =>
	conf -> ChunkSize -> NumCores -> arr a b -> arr [a] [b]
farmChunk conf chunkSize numCores f =
	unshuffle numCores >>>
	parEvalNLazy conf chunkSize (repeat (mapArr f)) >>>
	shuffle
\end{code}
\caption{Definition of |farmChunk|.}
\label{fig:farmChunk}
\end{figure}

@@

The |parEval2| skeleton is defined in Figure~\ref{fig:parEval2}. 
We start by transforming the |(a, c)| input into a two-element list |[Either a c]| by first tagging the two inputs with |Left| and |Right| and wrapping the right element in a singleton list with |return| so that we can combine them with |arr (uncurry (:))|. Next, we feed this list into a parallel Arrow running on two instances of |f +++ g| as described in the paper. After the calculation is finished, we convert the resulting |[Either b d]| into |([b], [d])| with |arr partitionEithers|. The two lists in this tuple contain only one element each by construction, so we can finally just convert the tuple to |(b, d)| in the last step.
\begin{figure}[h]
\begin{code}
parEval2 :: (ArrowChoice arr,
	ArrowParallel arr (Either a c) (Either b d) conf) =>
	conf -> arr a b -> arr c d -> arr (a, c) (b, d)
parEval2 conf f g =
	arr Left *** (arr Right >>> arr return) >>>
	arr (uncurry (:)) >>>
	parEvalN conf (replicate 2 (f +++ g)) >>>
	arr partitionEithers >>>
	arr head *** arr head
\end{code}
	\caption{Definition of |parEval2|.}
	\label{fig:parEval2}
\end{figure}
Furthermore, Fig.~\ref{fig:torus_example_rest} contains the omitted definitions of |prMMTr| (which calculates $AB^T$ for two matrices $A$ and $B$), |splitMatrix| (which splits the a matrix into chunks), and lastly |matAdd|, that calculates $A + B$ for two matrices $A$ and $B$.
\begin{figure}[h]
\begin{code}
prMMTr m1 m2 = [[sum (zipWith (*) row col) | col <- m2 ] | row <- m1]

@@

matAdd = chunksOf (dimX x) $ zipWith (+) (concat x) (concat y)
\end{code}
	\caption{Definition of |prMMTr|, |splitMatrix| and |matAdd|.}
	\label{fig:torus_example_rest}
\end{figure}

>>>>

The |mapArr| combinator (Fig.~\ref{fig:mapArr}) lifts any Arrow |arr a
b| to an Arrow |arr [a] [b]|. The original inspiration was from \citet{Hughes2005},
but the definition as then unified with |evalN|.

@@

instance Future BasicFuture a () where
    put _ = put'
    get _ = get'
\end{code}
\caption{|BasicFuture| type and its |Future| instance for the |Par| Monad and GpH Haskell backends.}
\label{fig:BasicFuture}
\end{figure}

@@

\begin{figure}[thb]
%parMap
\includegraphics[scale=0.7]{images/parMap}
\caption{|parMap| depiction.}
\label{fig:parMapImg}

@@

%parMapStream
\includegraphics[scale=0.7]{images/parMapStream}
\caption{|parMapStream| depiction.}
\label{fig:parMapStreamImg}

\begin{code}
parMapStream :: (ArrowParallel arr a b conf, ArrowChoice arr, ArrowApply arr) =>
	conf -> ChunkSize -> arr a b -> arr [a] [b]
parMapStream conf chunkSize f = parEvalNLazy conf chunkSize (repeat f)
\end{code}
\caption{|parMapStream| definition.}
\label{fig:parMapStream}
\end{figure}

@@

takeEach :: Int -> [a] -> [a]
takeEach n [] = []
takeEach n (x:xs) = x : takeEach n (drop (n-1) xs)
\end{code}
\caption{|shuffle|, |unshuffle|, |takeEach| definition.}
\label{fig:edenshuffleetc}
\end{figure}

@@

rightRotate :: (Arrow arr) => arr [a] [a]
rightRotate = arr $ \list -> case list of
  [] -> []
  xs -> last xs : init xs
\end{code}
\caption{|lazy| and |rightRotate| definitions.}
\label{fig:edenlazyrightrotate}
\end{figure}

@@

threetotwo :: (Arrow arr) => arr (a, b, c) (a, (b, c))
threetotwo = arr $ \ ~(a, b, c) -> (a, (b, c))
\end{code}
\caption{|lazyzip3|, |uncurry3| and |threetotwo| definitions.}
\label{fig:lazyzip3etc}
\end{figure}

\begin{figure}[h]
\begin{code}
farmChunk :: (ArrowParallel arr a b conf, ArrowParallel arr [a] [b] conf, 
             ArrowChoice arr, ArrowApply arr) =>
	conf -> ChunkSize -> NumCores -> arr a b -> arr [a] [b]
farmChunk conf chunkSize numCores f =
	unshuffle numCores >>>
	parEvalNLazy conf chunkSize (repeat (mapArr f)) >>>
	shuffle
\end{code}
\caption{|farmChunk| definition.}
\label{fig:farmChunk}
\end{figure}

@@

The |parEval2| skeleton is defined in Figure~\ref{fig:parEval2}. 
We start by transforming the |(a, c)| input into a two-element list |[Either a c]| by first tagging the two inputs with |Left| and |Right| and wrapping the right element in a singleton list with |return| so that we can combine them with |arr (uncurry (:))|. Next, we feed this list into a parallel Arrow running on two instances of |f +++ g| as described in the paper. After the calculation is finished, we convert the resulting |[Either b d]| into |([b], [d])| with |arr partitionEithers|. The two lists in this tuple contain only one element each by construction, so we can finally just convert the tuple to |(b, d)| in the last step.
\begin{figure}[h]
\begin{code}
parEval2 :: (ArrowChoice arr,
	ArrowParallel arr (Either a c) (Either b d) conf) =>
	conf -> arr a b -> arr c d -> arr (a, c) (b, d)
parEval2 conf f g =
	arr Left *** (arr Right >>> arr return) >>>
	arr (uncurry (:)) >>>
	parEvalN conf (replicate 2 (f +++ g)) >>>
	arr partitionEithers >>>
	arr head *** arr head
\end{code}
	\caption{|parEval2| definition.}
	\label{fig:parEval2}
\end{figure}
Furthermore, Fig.~\ref{fig:torus_example_rest} contains the omitted definitions of |prMMTr| (which calculates $AB^T$ for two matrices $A$ and $B$), |splitMatrix| (which splits the a matrix into chunks), and lastly |matAdd|, that calculates $A + B$ for two matrices $A$ and $B$.
\begin{figure}[h]
\begin{code}
prMMTr m1 m2 = [[sum (zipWith (*) row col) | col <- m2 ] | row <- m1]

@@

matAdd = chunksOf (dimX x) $ zipWith (+) (concat x) (concat y)
\end{code}
	\caption{|prMMTr|, |splitMatrix| and |matAdd| definition.}
	\label{fig:torus_example_rest}
\end{figure}
